---
title: "Olive data"
subtitle: "Ordering displays"
fontsize: 9pt
geometry: margin=.75in
output:
  pdf_document: default
  html_document:
    df_print: paged
header-includes:
- \usepackage{graphicx}
- \usepackage{color}
graphics: yes
classoption: letter
---

**25 marks**


This data are easily available from the `loon` package:


```{r, tidy=TRUE}
library(loon) 
# The first three rows of which are
head(olive, 3)
```

a. **(2 marks)** Separate the data on `oleic` into 9 different groups as defined by the olive growing `Area`, and draw side by side boxplots of all 9 groups.  
Colour the boxplots uniquely using
    ```{r, echo=TRUE, message=FALSE, error = FALSE, warning=FALSE}
    library(colorspace)
    cols <- rainbow_hcl(9) # Use these colours.
    
    data = split(olive$oleic, olive$Area)
    names(cols)=names(data)
    par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)
    boxplot(data, col = cols, xaxt='n', main="Olive oleic at different Areas", ylab="oleic")
    legend("topright", legend = names(cols), col=cols, lty=1,cex=0.8, inset=c(-0.4,0))
    ```
    
    Show your code together with your output.
    
b. Load the `R` package `PairViz` (i.e. `library(PairViz)`).  Use the variate `oleic` and the same colours for the olive growing areas as in part (a) throughout the following:
    i. **(3 marks)** Suppose we wish every pair of boxplots to appear next to one another in the same plot.  
        - How many such pairwise comparisons exist?
        - Give the code that will construct this display (without any other constraint on the ordering).
        - Show the display which resulted from your code. 
        
    ```{r}
    library(PairViz)
    ord = eulerian(9)
    par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)
    boxplot(data[ord], col = cols[ord],xaxt="n", main="Olive oleic at different Areas", ylab="oleic")
    legend("topright", legend = names(cols), col=cols, lty=1,cex=0.8, inset=c(-0.4,0))
    ```
    
number of pairwise that exists is $\left(^9_2 \right)=36$
        
    ii. **(5 marks)** Suppose we wish every pair of boxplots to appear next to one another in the same plot but that the boxplots should be grouped so that all areas appear only once in each group.  
        - Maintaining the same colours for the areas as before, give the code that will construct this display (without any other constraint on the ordering).  
        - Show the display which resulted from your code. 
```{r}
ordHam = hpaths(9)
for (i in 1:4){
  par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)
  boxplot(data[ordHam[i,]], col = cols[ordHam[i,]],xaxt="n", main="Olive oleic at different Areas", ylab="oleic")
  legend("topright", legend = names(cols), col=cols, lty=1,cex=0.8, inset=c(-0.4,0))
}
```
        
    iii. **(7 marks)** Construct $t$ tests for every pair of olive growing areas (recall  `pairwise.t.test` from class).  Use the significance levels from these tests to construct an ordering of the boxplot pairs, one which favours having the most significantly different pairs at the left of the display.  
        - Show your code.
        - Show the resulting display.
        - Does the ordering perfectly arrange the boxplots so that for any pairwise comparison, those to the left are more significant and those to the right are less significant? 
        - Explain why the ordering was successful (or unsuccessful) in this way.
        - Show a display showing only the first 8 comparisons.

```{r}
    test = pairwise.t.test(olive$oleic, olive$Area)
    pvals = test$p.value
    
    Area.names = names(data)
    narea = length(Area.names)
    
    oleics = matrix(0, nrow=narea, ncol=narea)
    rownames(oleics) = Area.names
    colnames(oleics) = Area.names
    oleics[2:narea, 1:(narea-1)] = pvals 
    
    
    diag(oleics) <- 0
    for(i in 1:(narea -1)) {
      for(j in (i+1):narea) {
        oleics[i,j] <- oleics[j,i]
      }
    }
    
    low2high = eulerian(oleics)
    par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)
    boxplot(data[low2high],col =  cols[low2high], xaxt='n', main="Olive oleic at different Areas", ylab="oleic")
    legend("topright", legend = names(cols), col=cols, lty=1,cex=0.8, inset=c(-0.4,0))


    par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)
    boxplot(data[low2high[1:8]], col = cols[low2high[1:8]], xaxt='n', main="Olive oleic at different Areas", ylab="oleic")
    legend("topright", legend = names(cols), col=cols, lty=1,cex=0.8, inset=c(-0.4,0))
    
    
    
```

Does the ordering perfectly arrange the boxplots so that for any pairwise comparison, those to the left are more significant and those to the right are less significant? 

No.

Explain why the ordering was successful (or unsuccessful) in this way.

Denote T score for $H_0$: area A and area B has the same oleic by $T_{AB}$

If smallest t-score is $T_{AB}$ AND second smallest t-score is $T_{CD}$

Then, there is no way to put (boxplot of A or B) next to (boxplot of C or D) if ABCD are different Areas(which is entirely possible).
Then what we actually put on the right of $T_{AB}$ would be larger than $T_{CD}$, breaking the rule that comparison on the left are more significant.
        
c. The olive growing areas are divided into three different regions: North, South, and Sardinia.  In this part of the question, interest lies only in comparisons between each growing area in the south  and each area in Sardinia.  That is, each southern area (4 areas) is to be compared to each Sardinian area  (2 areas) yielding a total of 8 comparisons of interest.
    
    i. **(4 marks)**  Having loaded `PairViz`, create a graph having all six areas in the South and Sardinia as nodes and with edges between every pair whose comparison is of interest.
        - plot this graph
        - show the code used to create the graph and to plot it.
```{r}
    sdata = split(olive, olive$Region)
    sdata.south = sdata$South
    south.nodes = split(sdata.south$oleic, sdata.south$Area,drop = TRUE)
    sdata.Sardinia= sdata$Sardinia
    sardinia.nodes = split(sdata.Sardinia$oleic, sdata.Sardinia$Area,drop = TRUE)
    p = bipartite_graph(names(sardinia.nodes), names(south.nodes))
    plot(p)
```
            
    ii. **(4 marks)**  Using the graph from part (i), construct an Eulerian and use that Eulerian to produce a sequence of boxplots that show the comparisons of interest.
        - show the boxplot display
        - show the code used to construct the Eulerian and the display.
           
```{r}
ord = eulerian(p) 
par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)
boxplot(data[ord], col=cols[ord], xaxt='n', main="Olive oleic at different Areas", ylab="oleic")
legend("topright", legend = names(cols), col=cols, lty=1,cex=0.8, inset=c(-0.4,0))
``` 